Quantum-Classical Phase Transition for Exciton-Polariton Josephson Junctions

### 2017-02-14 writing/phase_transition/phase_transition.tex
Dissipation: outset

Another type of barrier --- upper barrier.
Exist when $\Lambda > 2(1 - \lambda)$.
Potential presumes local minima at $\pm 2K(\lambda)$ and local maxima at $z_{max} = \pm \cn^{-1} \left( -\frac{2(1 - \lambda)}{\Lambda} \right)$.
Equation for the thermon yeilds small ocsilation frequency near the bottom $z_{max}$ of the inverted potential $\omega_0 = \alpha s \sqrt{\frac{\Lambda^2}{1 - \lambda} - 4(1 - \lambda)}$.
The crossover temperature of second-order phase transition
%
\begin{equation}
T_{c}^{(2)} = (\hbar \alpha s / 2 \pi k_B) \sqrt{\frac{\Lambda^2}{1 - \lambda} - 4(1 - \lambda)}.
\label{eq:second_order_2}
\end{equation}
%
With the aim of analytical determining of the phase transition order let's write the Taylor series for $V(z)$ near the vicinity of $z_{max}$.
Expansion of the potential $V(z)$ excluding constant value $V(z_{max})$ gives:
%
\begin{equation}
V(\tilde{z}) = \alpha s^2 \Big[ -\tfrac{1}{4} \left( \tfrac{\Lambda^2}{1 - \lambda} - 4(1 - \lambda) \right) \tilde{z}^2 
+ \tfrac{1}{2} \tfrac{\Lambda \sqrt{\Lambda^2 - 4(1 - \lambda)^2}}{\sqrt{1 - \lambda} \sqrt{\Lambda^2 + 4 \lambda (1 - \lambda}} \tilde{z}^3 + o(\tilde{z}^4) \Big],
\label{eq:potential_teylor_2}
\end{equation}
%
where $\tilde{z} = z - z_{max}$.
It follows from (\ref{eq:potential_teylor_2}) that $-V(\tilde{z})$ aways has the form $\tilde{z}^2 - \tilde{z}^3$, as $\Lambda > 0$, so in this case only second-order phase transition occurs.
We plot typical behaviour of the $\tau(E)$, $S(T)$ and $T_{c}^{(2)}$ on Figs. \ref{pic:action_period_temperature_2}, and summarize our results on the Fig. \ref{pic:phase_boundary_combined}.
%
\begin{figure}[ht]
\begin{minipage}[h]{0.49\linewidth}
\center{\includegraphics[width=\linewidth]{pic/action_period_2nd_order_transition_b2.eps} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}[h]{0.49\linewidth}
\center{\includegraphics[width=\linewidth]{pic/phase_transition_temperatures_b2.eps} \\ (b)}
\end{minipage}
\caption{\dots \label{pic:action_period_temperature_2}}
\end{figure}
%
\section{Phase transitions in the presence of non-equilibrium processes 
\label{sec:phase_transition}}

\dots

>>>>

Another type of barrier --- upper barrier.
Exist when $\Lambda > 2(1 - \lambda)$.
Potential presumes local minima at $\pm 2K(\lambda)$ and local maxima at $z_{max} = \pm \cn^{-1} \left( -\frac{2(1 - \lambda)}{\Lambda} \right)$.
Equation for the thermon yeilds small ocsilation frequency near the bottom $z_{max}$ of the inverted potential $\omega_0 = \alpha s \sqrt{\frac{\Lambda^2}{1 - \lambda} - 4(1 - \lambda)}$.
The crossover temperature of second-order phase transition
%
\begin{equation}
T_{c}^{(2)} = (\hbar \alpha s / 2 \pi k_B) \sqrt{\frac{\Lambda^2}{1 - \lambda} - 4(1 - \lambda)}.
\label{eq:second_order_2}
\end{equation}
%
With the aim of analytical determining of the phase transition order let's write the Taylor series for $V(z)$ near the vicinity of $z_{max}$.
Expansion of the potential $V(z)$ excluding constant value $V(z_{max})$ gives:
%
\begin{equation}
V(\tilde{z}) = \alpha s^2 \Big[ -\tfrac{1}{4} \left( \tfrac{\Lambda^2}{1 - \lambda} - 4(1 - \lambda) \right) \tilde{z}^2 
+ \tfrac{1}{2} \tfrac{\Lambda \sqrt{\Lambda^2 - 4(1 - \lambda)^2}}{\sqrt{1 - \lambda} \sqrt{\Lambda^2 + 4 \lambda (1 - \lambda}} \tilde{z}^3 + o(\tilde{z}^4) \Big],
\label{eq:potential_teylor_2}
\end{equation}
%
where $\tilde{z} = z - z_{max}$.
It follows from (\ref{eq:potential_teylor_2}) that $-V(\tilde{z})$ aways has the form $\tilde{z}^2 - \tilde{z}^3$, as $\Lambda > 0$, so in this case only second-order phase transition occurs.
We plot typical behaviour of the $\tau(E)$, $S(T)$ and $T_{c}^{(2)}$ on Figs. \ref{pic:action_period_temperature_2}, and summarize our results on the Fig. \ref{pic:phase_boundary_combined}.
%
\begin{figure}[ht]
\begin{minipage}[h]{0.49\linewidth}
\center{\includegraphics[width=\linewidth]{pic/action_period_2nd_order_transition_b2.eps} \\ (a)}
\end{minipage}
\hfill
\begin{minipage}[h]{0.49\linewidth}
\center{\includegraphics[width=\linewidth]{pic/phase_transition_temperatures_b2.eps} \\ (b)}
\end{minipage}
\caption{\dots \label{pic:action_period_temperature_2}}
\end{figure}
%
\section{Phase transitions in the presence of non-equilibrium processes 
\label{sec:non-equilibrium}}

Consider the influence of the dissipation in our system.
Supposing that $\Gamma = 0$, system with dissipation can be derived from hamiltonian and reads as
%
\begin{equation}
\begin{array}{lcl}
i \psi_{1, t} & = & -i \kappa \psi_1 +  \alpha |\psi_1|^2 \psi_1 - (B/2 - \beta \psi_1^* \psi_2) \psi_2; \\
i \psi_{2, t} & = & -i \kappa \psi_2 + \alpha |\psi_2|^2 \psi_2 - (B/2 - \beta \psi_2^* \psi_1) \psi_1.
\end{array}
\end{equation}
%
Change of variables: $\psi_{1,2} = \Psi_{1,2} \exp(-\kappa t)$.
System takes form
%
\begin{equation}
\begin{array}{lcl}
i \Psi_{1, t} & = & \alpha(t) |\Psi_1|^2 \Psi_1 - (B/2 - \beta(t) \Psi_1^* \Psi_2) \Psi_2; \\
i \Psi_{2, t} & = & \alpha(t) |\Psi_2|^2 \Psi_2 - (B/2 - \beta(t) \Psi_2^* \Psi_1) \Psi_1.
\end{array}
\end{equation}
%
It's equivalent to the non-dissipative system with time dependent parameters $\alpha(t) = \alpha \exp(-2 \kappa t)$, $\beta(t) = \beta \exp(-2 \kappa t)$.
Hamiltoninan in terms of Stocks operators (\ref{eq:pseudo_spin}):
%
\begin{equation}
\hat{H} = \alpha(t) \hat{S}_z^2 + \beta(t) \hat{S}_x^2 - B \hat{S}_x.
\end{equation}
%
Equaitions on the mean-values of operators:
% 
\begin{equation}
\begin{array}{lcl}
\dot{S}_z & = &  2 \beta(t) S_x S_y - B S_y; \\
\dot{S}_x & = & -2 \alpha(t) S_z S_y; \\
\dot{S}_y & = & 2(\alpha(t) - \beta(t)) S_z S_z + B S_z.
\end{array}
\end{equation}
%
Invariant of the system
%
\begin{equation}
S_x^2 + S_y^2 + S_z^2 = S_0^2 = R^2,
\end{equation}
%
where $R$ -- is a radius of the Poincare sphere.
Non-dissipative system has a bifurcation at the critical values of $\beta^* = B / 2R$.

It can be transformed to the equation on the phase difference $\phi$ and $S_z$ as follows
%
\begin{equation}
\begin{array}{lcl}
\dot{S}_z & = & B \sqrt{1 - S_z^2} \sin \phi - 2 \beta(t) (1 - S_z^2) \sin \phi \cos \phi; \\[10pt]
\dot{\phi} & = &  -2 \alpha(t) S_z + 2 \beta(t) S_z \cos^2 \phi - \dfrac{B S_z \cos \phi}{\sqrt{1 - S_z^2}}.
\end{array}
\end{equation}
%
We studied its dynamic numerically.
%
\begin{figure}[ht]
\center{\includegraphics[width=0.7\linewidth]{pic/phase_potential_dissipation.eps}}
\caption{\dots}
\label{pic:phase}
\end{figure}
%
Non-dissipative system hase two different regimes depend on different values of $B$: blue and red.
Case blue ($\beta > \beta^*$) corresponds to the existence of barrier, and the red case ($\beta < \beta^*$) corresponds to the absence of the barrier.

Linearization ($z \ll 1$, $\phi \ll 1$) of the system:
%
\begin{equation}
\begin{array}{lcl}
\dot{z} & = & (B - 2\beta(t)) \phi; \\
\dot{\phi} & = & (2 \beta(t) - 2 \alpha(t) - B) z.
\end{array}
\end{equation}
%
Adiabaticity condition:
%
\begin{equation}
\dfrac{1}{\omega^2(t)} \Big| \dfrac{d \omega}{d t} \Big| \ll 1.
\end{equation}
%
It allows us to reduce linear system to the harmonic oscillator equaiton with slowly evolving frequency:
%
\begin{equation}
\ddot{\phi} + \omega^2(t) \phi = 0,
\end{equation}
%
$\omega(t) = (B - 2\beta(t))(B - 2 \beta(t) + 2 \alpha(t))$.
We can approximate its solution as
%
\begin{equation}
\phi(\tau) \approx A / \sqrt{\omega(\tau)} \cos (\omega(\tau) \tau),
\end{equation}
%
where $\tau$ is the time elapsed after the oscillation sets in, and $A$ is an amplitude of oscillations at the moment $\tau = 0$.
